State and Input Dependent Disturbances in Min-Max Model Predictive Control

### 2015-11-12 draft.tex
edits to introduction and references

\author{Rainer~M.~Schaich,~\IEEEmembership{Student Member,~IEEE,}
        Mark~Cannon
\thanks{The authors are with the Department of Engineering Science, University of
Oxford, Parks Road, Oxford, OX1~3PJ, e-mail: rainer.schaich@eng.ox.ac.uk, 
mark.cannon@eng.ox.ac.uk}% <-this % stops a space
\thanks{Manuscript received April 19, 2005; revised September 17, 2014.}}

@@

Robust model predictive control (RMPC) methods compute control laws for uncertain systems that optimise a 
performance criterion for all possible realisations of the uncertainties.
%
One major challenge in RMPC is to guarantee feasibility of the problem for all possible uncertainties in 
presence of state constraints.
%
This can be achieved by introducing specifically designed constraints in such a way that at the
end of the control horizon the system state is contained in a set that is feasible and invariant for all
possible realisations of uncertainty and for a given feedback controller, in particular the largest such set, 
the so called \emph{maximal robust positive invariant} (MRPI) set.
%
Although this concept was introduced shortly after Witsenhausen proposed the use of a game theoretic approach
to robust model predictive control in~\cite{Witsenhausen:1968} by several authors~\cite{Bertsekas:1971,Glover:1971},
it took over two decades before the first MRPI sets could be algorithmically determined explicitly.
%
The main contribution of this paper is to discuss the use of parametrised sets to model uncertainties, as a parameter
we use the state and the input allowing us to model linearisation errors and multiplicative uncertainties in an 
additive way.
%
We develop the abstract and computational tools necessary to characterise and determine these sets, we introduce the 
concept of \emph{parametric convexity} which is essential in order to obtain convex MRPI and controllable sets.

The algorithmic solution of RMPC problems suffered a similar history as the MRPI sets did,

Throughout this paper we denote by $f\vert_{\mathcal C} = \{y:\exists x\in\mathcal C\;y=f(x)\}$
the image of $\mathcal C$ under the map $f$, for polytopes defined by intersection of the halfspaces 
$a_i x \leq b_i$ for $i\in\{1,\dots,m\}$ we denote the matrix~${\bf{a}}_{\mathcal I}$ and vector~
${\bf{b}}_{\mathcal I}$ with the rows $a_i$ and $b_i$ for all indices in the index set $i\in\mathcal 
I\subseteq\{1,\dots,m\}$.

\section{Problem statement}

@@

for elaborate discussions) can be solved efficiently with an active-set solver (see e.g.~\cite{Fletcher:1987}).

>>>>

\author{Rainer~M.~Schaich,~\IEEEmembership{Student Member,~IEEE,}
        Mark~Cannon
\thanks{The authors are with the Department of Engineering Science, University of
Oxford, Parks Road, Oxford, OX1~3PJ, e-mail: rainer.schaich@eng.ox.ac.uk, 
mark.cannon@eng.ox.ac.uk}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised September 17, 2014.}
}

@@

Robust model predictive control (RMPC) is a control strategy for uncertain systems that optimises a 
performance objective for all possible realisations of model uncertainty.
%
First introduced in~\cite{Witsenhausen:1968}, the min-max RMPC problem is formulated as a sequence of games in which an adversary maximises the effect of model disturbances on a performance objective, and the controller solves the optimal
control problem for the resulting worst-case system.
%
We discuss the \emph{closed-loop setup} where, at each stage of a prediction horizon, the disturbance input maximizes the sum of the stage cost and the cost-to-go, and the control input minimises this cost so that the controller reacts to the worst-case disturbance at each stage.
%
The closed-loop setup includes, as a special case, the \emph{open-loop setup} in which the disturbance is allowed to perturb the 
entire trajectory before the controller is allowed to react, since this can be understood as a closed loop setup (with a horizon length of one) applied to a lifted system.
%

A major challenge in RMPC is to guarantee feasibility of the control problem for all possible uncertainties in 
presence of state constraints.
%
This can be achieved by introducing additional constraints into the problem that force the state at the end of the horizon to lie in a set that is feasible and invariant for all possible realisations of uncertainty under a given feedback law; the largest such set, 
the so-called \emph{maximal robust positively invariant} (MRPI) set, is often used for this purpose.
%
The concept of MRPI sets and their analytical properties were introduced in the literature (e.g.~\cite{Bertsekas:1971,Glover:1971}) shortly after robust model predictive control was originally proposed. However algorithms for numerically 
computing MRPI sets were not developed until more than two decades later~\cite{Blanchini:1994,DeSantis:1994,Kolmanovsky:1998};
earlier algorithms either failed to ensure finite determinability
of the MRPI set (see e.g.~\cite{Blanchini:1990}), 
or were not guaranteed to produce the maximal RPI set (e.g.~\cite{Blanchini:1991}).
%

This paper considers the use of parametrised sets for model uncertainties, which were recently introduced in~\cite{Schaich:CDC:2015,Schaich:NMPC:2015}. As parameters we consider the state and the control input; this allows linearisation errors and multiplicative uncertainties to be modelled as additive disturbances.
%
The paper's contribution is to develop
%extends~\cite{Schaich:CDC:2015} by developing 
the analytical and computational tools that are needed to characterise and determine invariant and controllable sets for systems with parametrised uncertainty sets, and to implement them in a closed loop RMPC strategy. To this end, we provide necessary and sufficient conditions for the property, introduced in~\cite{Schaich:CDC:2015}, of \emph{parametric convexity}, which is essential in order to obtain convex MRPI and controllable sets, and we derive a bound on the number of iterations needed to compute the MRPI set. The current paper also develops and extends the min-max RMPC approach introduced in~\cite{Schaich:NMPC:2015} by considering the case of degenerate constraints.

To solve RMPC problems a few different methods are available. For the case of unknown additive disturbances, 
a scenario-based approach is proposed in~\cite{Scokaert:1998} for the online solution of min-max RMPC problems. However, the computational complexity of this approach grows exponentially with the length of the horizon. 
%
A parametrisation of predicted feedback laws proposed in~\cite{Rakovic:2012} reduces the RMPC problem to a single linear 
program to be solved online, but although this is a closed loop strategy the parametrisation is in general 
suboptimal. 
%
Explicit RMPC algorithms (e.g.~\cite{Bemporad:2003,Diehl:2004}) optimise the worst case predicted performance for each admissible model state offline and
use results from multi-parametric programming to parametrise the optimal control law as a piecewise affine function of the
state.
%
However, this requires extensive offline computation and becomes impractical for a high-dimensional state space.
%
In this paper we use an active-set approach based on that presented in~\cite{Buerger:ACC,Buerger:IJRNC,Schaich:NMPC:2015}.
In a similar manner to explicit RMPC algorithms, we use multi-parametric programming to obtain solutions efficiently
for a given active-set, however, we avoid an exploration of the entire feasible set of states but rather solve online for the current state only.
%
Our approach uses a line search to update the set of active constraints, thus generating
the solution of the RMPC problem online in an efficient manner.

The paper is structured as follows, in section~\ref{sec:problem:statement} we present the RMPC problem under consideration
and define the constraint sets for the model uncertainty in terms of state and input dependent disturbance sets.
%
The properties of parametrised sets are then considered in section~\ref{sec:p:convex:sets}, in particular we discuss
parametric convexity and its characterisation in section~\ref{ssec:properties:of:p:convex:maps}, and its use in the computation of MRPI sets, which is presented in section~\ref{ssec:MRPI:sets}.
%
To illustrate how state and input dependent disturbances can be used to account for nonlinearities and 
calculate the MRPI set we present an example in section~\ref{sec:example:I}.
%
The solver we present here relies entirely on a recursive use of multi-parametric programming techniques, 
which are presented in section~\ref{ssec:recursive:mpQP} for a given active-set, the active set is then
updated using a line search which is described in subsection~\ref{ssec:line:search}. In order to account for the 
degeneracy that can occur in the optimisation, we consider the case of degenerate problems in subsection~\ref{ssec:degeneracy}.
%
In section~\ref{sec:solving:min:max:programs} we translate the general formulation of section~\ref{sec:recursive:mpQP:via:line:search}
to the problem presented in section~\ref{sec:problem:statement}, and section~\ref{sec:complexity}
deals with the complexity of the proposed algorithm.
%
We revisit the example system presented in section~\ref{sec:example:I} to illustrate the proposed RMPC
algorithm in section~\ref{sec:example:II}.
%
Section~\ref{sec:conclusion} provides brief conclusions.

\emph{Notation:} Throughout this paper we denote 
the image of a set $\mathcal C\subseteq X$ under the map $f:X\mapsto Y$ as $f\vert_{\mathcal C} = \{y\in Y:\exists x\in\mathcal C,\;y=f(x)\}$. 
%
Given a collection of row-vectors $a_i$ and scalars $b_i$ for $i\in\{1,\ldots,m\}$, we denote as ${\bf{a}}_{\mathcal I}$ and
${\bf{b}}_{\mathcal I}$ the matrix and vector, respectively, with rows $a_i$ and $b_i$ for all indices $i$ in the index set $\mathcal I\subseteq\{1,\dots,m\}$.

%To represent the facets of a polytope defined by intersection of the half spaces $\{x : a_i x \leq b_i\}$ for $i\in\{1,\dots,m\}$, we denote as ${\bf{a}}_{\mathcal I}$ and ${\bf{b}}_{\mathcal I}$ the matrix and vector, respectively, with rows $a_i$ and $b_i$ for all indices in the index set $i\in\mathcal I\subseteq\{1,\dots,m\}$.

\section{Problem statement}\label{sec:problem:statement}

@@

for elaborate discussions) can be solved efficiently with an active-set solver (see e.g.~\cite{Fletcher:2000}).
